Learning semi-Markovian DAGs with flow-based VAE

### 2023-12-11 Tex/slide.tex
paper, slide DAG learning 추가

\documentclass{beamer}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{kotex}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage[backend=bibtex, style=authortitle-comp]{biblatex}
\addbibresource{references.bib}

@@

\begin{frame}{DAG and SEM}
    \begin{itemize}
        \item Directed Acyclic Graph(DAG) $\mathcal G = (V, E)$ with $V = \{1, \dots, m\}$ is used to represent a causal structure \footcite{pearl_2009}
        \item Linear Structural Equation Model(SEM) is defined as
        \begin{equation*}
            X_i = \sum_{j \in \text{Pa}(i)} \beta_{ij} X_j + \epsilon_i
        \end{equation*}
        where $\epsilon_i$ is a noise variable.
        \item Matrix form can be written as
        \begin{equation}
            \mathbf X = \mathbf A^T \mathbf X + \mathbf \epsilon
        \end{equation}
        where $\mathbf A$ is an adjacency matrix of $\mathcal G$ and $\mathbf \epsilon$ is a noise vector.
    \end{itemize}
\end{frame}

\begin{frame}{DAG and SEM(Cont'd)}
    \begin{itemize}
        \item If $(\mathbf I-\mathbf A^T)$ is nonsingular, SEM can be written as
        \begin{equation}
            \mathbf X = (\mathbf I - \mathbf A^T)^{-1} \mathbf \epsilon
        \end{equation}
        \item Usually we assume $\mathbf \epsilon \sim \mathcal N(\mathbf 0, \mathbf \Sigma)$ and
        \begin{equation*}
            \mathbf X \sim \mathcal N(\mathbf 0, (\mathbf I - \mathbf A^T)^{-1} \mathbf \Sigma (\mathbf I - \mathbf A)^{-1})
        \end{equation*}
    \end{itemize}
\end{frame}

\begin{frame}{Semi-Markovian DAG}
    \begin{itemize}
        \item Semi-Markovian DAG is a DAG where the noise variables have dependencies \footcite{Shipster2006semimarkovian}
        \item The noise vector $\mathbf \epsilon$ is a semi-Markovian noise vector if
        \begin{equation}
            \Sigma \in \mathcal{W}(G) := \{ \Sigma \in \mathbb R^{m\times m} : \Sigma_{ij} = \Sigma_{ji} = 0 \text{ if } i \notin \text{sib}(j) \}
        \end{equation}
        where $\text{sib}(j) = \{i : i \sim j \text{ in } G\}$ \footcite{wang2017empirical}
        \item Learn cyclic adjacency matrix or acyclic adjacency matrix
    \end{itemize}
\end{frame}

\begin{frame}{DAG Learning}
    \begin{itemize}
        \item DAG learning is a problem of estimating $\mathcal G$ from $\mathbf X$
        \item NP-hard problem \footcite{ChickeringNPhard}
        \item Score-based methods \footcite{Spirtes2000} and constraint-based methods \footcite{ChickeringGES} are used
        \item NOTEARS\footcite{zheng2018dags}: Combinatorial optimization to Continuous optimization problem
        \begin{equation*}
            \min_{\mathbf{A}} F(A) \text{ subject to } G(A) \in \mathbb{D} \Rightarrow 
            \min_{\mathbf{A}} F(A) \text{ subject to } h(A)=0
        \end{equation*}
        where $\mathbb D$ denotes the discrete space of DAGs on $m$ nodes and $h(A)=0$ iff $G(A)$ is acyclic.
    \end{itemize}
\end{frame}

\begin{frame}{DAG Learning(Cont'd)}
    \begin{block}{Theorem}
        A matrix $\mathbf A$ is the adjacency matrix of a DAG if and only if
        \begin{equation*}
            h(A) = \rm{tr}(\exp(\mathbf A \circ \mathbf A)) = m
        \end{equation*}
        where $\circ$ denotes the Hadamard product.
    \end{block}
    \begin{block}{Theorem}
        A matrix $\mathbf A$ is the adjacency matrix of a DAG if and only if
        \begin{equation*}
            \rm{tr}[(I + \alpha \mathbf A \circ \mathbf A)^m] = m
        \end{equation*}
        for any $\alpha>0$.
    \end{block}
\end{frame}

\subsection*{VAE and Normalizing Flow}

\begin{frame}{VAE}
    \begin{itemize}
        \item Variational Autoencoder(VAE) is a generative model that learns a latent variable model
        \item It is trained by maximizing the evidence lower bound(ELBO)
        \begin{equation}
            \mathcal L(\theta, \phi; \mathbf x) = \mathbb E_{q_\phi(\mathbf z | \mathbf x)}[\log p_\theta(\mathbf x | \mathbf z)] - \text{KL}(q_\phi(\mathbf z | \mathbf x) || p(\mathbf z))
        \end{equation}
        \item $q_\phi(\mathbf z | \mathbf x)$ is an approximate posterior and $p_\theta(\mathbf x | \mathbf z)$ is a generative model
    \end{itemize}
\end{frame}

\begin{frame}{Normalizing Flow}
    \begin{itemize}
        \item Normalizing flow transforms a simple distribution to a complex distribution \footcite{rezende15normalizingflows}
        \item It is a sequence of invertible transformations $f_t : \mathbb R^d \to \mathbb R^d$
        \begin{equation}
            z_T = f_T \circ f_{T-1} \circ \cdots \circ f_1(z_0)
        \end{equation}
        \item The probability density function of $z_T$ is
        \begin{equation}
            p(z_T) = p(z_0) \left| \det \left( \frac{\partial f_T \circ f_{T-1} \circ \cdots \circ f_1(z_0)}{\partial z_0} \right) \right|
        \end{equation}
    \end{itemize}
\end{frame}

\begin{frame}{Linear IAF}
    \begin{itemize}
        \item Inverse Autoregressive Flow(IAF) is a normalizing flow that uses autoregressive transformation \footcite{kingma2016iaf}
        \item Linear IAF is a special case of IAF where $f_t$ is a linear transformation
        \begin{equation}
            \bf z_t = \bf \mu_t + \bf \sigma_t \odot \bf z_{t-1}.
        \end{equation}
        \item Transformation of the linear IAF can be written as
        \begin{equation}
            \mathbf{z}_T = \mathbf{L(x)} \cdot \mathbf{z}_0.
        \end{equation}
        where $\mathbf{L(x)}$ is a lower triangular matrix from encoder network.
        \item We can use more flexible prior distribution by using normalizing flow.
        \item When $\mathbf z_0$ has diagonal covariance, $\mathbf z_T$ then has full covariance matrix.
    \end{itemize}
\end{frame}

@@

\begin{frame}{Learning}
    \begin{itemize}
        \item The optimization procedure is minimizing the following loss function:
        \begin{equation}
            \mathcal{L}(A, W, \lambda) =-\mathcal{L}_{\mathrm{ELBO}} + \tau \Vert A \Vert_1 + \lambda h(A) + \frac{c}{2} |h(A)|^2.
        \end{equation}
        \item The second term is the L1 regularization term, which encourages sparsity of the adjacency matrix.
        \item $\tau=0.1 \times ({m\over 50})^2$ works well in our experiment setting.
        \item The third and fourth terms are the augmented Lagrangian terms.
        \item Gradually increasing the value of $c$ during the training process results the acyclicity constraint to zero.
    \end{itemize}
\end{frame}

@@

\begin{frame}{Experiment}
    \begin{itemize}
        \item We compare our method with the DAG-GNN \footcite{yu2019daggnn} with random graph datasets.
        \item thresholding value of extracting graph as 0.3
        \item Erdős–Rényi random graph with 5000 samples
        \item Node size : 10, 20, 30, and 50 nodes
        \item Noise type : (Independent) Gaussian, Laplace, Exponential and (Dependent) Gaussian
    \end{itemize}
\end{frame}

\begin{frame}{Experiment(Cont'd)}
    \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{fig/mixed_graph_example.png}
        \caption{Example of mixed graph}
        \label{fig:mixed_graph_example}
    \end{figure}
\end{frame}

\begin{frame}{Experiment(Cont'd)}
    \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{fig/mixed_graph_adjacency.png}
        \caption{Adjacency matrix of previous example graph}
        \label{fig:mixed_graph_adjacency}
    \end{figure}
\end{frame}

@@

\begin{frame}{Independent Noise}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/SHD_independence_normal.pdf}
                \caption{SHD with normal}
                \label{fig:ind_gaussian_shd}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/FDR_independence_normal.pdf}
                \caption{FDR with normal}
                \label{fig:ind_gaussian_fdr}
            \end{figure}
        \end{column}
    \end{columns}
    \begin{itemize}
        \item In the case of independent Gaussian noise, DAG-GNN(orange) performs better than our method(blue).
    \end{itemize}
\end{frame}

\begin{frame}{Independent Noise(Cont'd)}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/SHD_independence_laplace.pdf}
                \caption{SHD with Laplace}
                \label{fig:ind_laplace_shd}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/FDR_independence_laplace.pdf}
                \caption{FDR with Laplace}
                \label{fig:ind_laplace_fdr}
            \end{figure}
        \end{column}
    \end{columns}
    In the case of independent Laplace noise, our method(blue) performs better than DAG-GNN(orange).
\end{frame}

\begin{frame}{Independent Noise(Cont'd)}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/SHD_independence_exponential.pdf}
                \caption{SHD with Exponential}
                \label{fig:ind_exponential_shd}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/FDR_independence_exponential.pdf}
                \caption{FDR with Exponential}
                \label{fig:ind_exponential_fdr}
            \end{figure}
        \end{column}
    \end{columns}
    In the case of independent Exponential noise, our method(blue) performs better than DAG-GNN(orange).
\end{frame}

\begin{frame}{Independent Noise(Cont'd)}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/Predicted Edges_independence_laplace.pdf}
                \caption{Number of predicted edges with Laplace}
                \label{fig:ind_laplace_edges}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/Predicted Edges_independence_exponential.pdf}
                \caption{Number of predicted edges with Exponential}
                \label{fig:ind_exponential_edges}
            \end{figure}
        \end{column}
    \end{columns}
    Considering the number of predicted edges with the SHD and FDR, our method(blue) performs better than DAG-GNN(orange).
\end{frame}

\begin{frame}
    \begin{figure}
        \centering
        \includegraphics[height=3.5cm]{fig/comparison_indep_30_laplace_seed301.pdf}
        \label{fig:ind_exponential_comparison}
    \end{figure}
    \begin{figure}
        \centering
        \includegraphics[height=3.5cm]{fig/comparison_indep_30_exponential_seed301.pdf}
        \label{fig:ind_exponential_comparison}
        \captionsetup{justification=centering}
        \caption{Comparison of estimated graphs with Laplace(above) and Exponential(below) noise}
    \end{figure}
\end{frame}

\begin{frame}{Dependent Noise}
    Proportion of bidirectional edges fixed to 0.3.
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/SHD_dependence_30_DAG_threshold0.3.pdf}
                \caption{SHD with dependent Gaussian}
                \label{fig:dep_gaussian_shd}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/FDR_dependence_30_DAG_threshold0.3.pdf}
                \caption{FDR with dependent Gaussian}
                \label{fig:dep_gaussian_fdr}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}{Dependent Noise(Cont'd)}
    Proportion of bidirectional edges fixed to 0.3.
    \begin{figure}
        \centering
        \includegraphics[height=4cm]{fig/Predicted Edges_dependence_30_DAG_threshold0.3.pdf}
        \caption{Number of predicted edges with dependent Gaussian}
        \label{fig:dep_gaussian_edges}
    \end{figure}
    \begin{itemize}
        \item In the case of dependent Gaussian noise, our method(blue) performs better than DAG-GNN(orange).
        \item Considering the stable number of predicted edges with the SHD and FDR, our method(blue) outperforms.
    \end{itemize}
\end{frame}

\begin{frame}{Dependent Noise(Cont'd)}
    Node size $m=10$ : similar performance between DAG-GNN(orange) and our method(blue).
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/SHD_dependence_10_DAG_threshold0.3.pdf}
                \caption{SHD with dependent Gaussian}
                \label{fig:dep_gaussian_shd_10}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/FDR_dependence_10_DAG_threshold0.3.pdf}
                \caption{FDR with dependent Gaussian}
                \label{fig:dep_gaussian_fdr_10}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}{Dependent Noise(Cont'd)}
    Node size $m=20$ : our method(blue) performs better than DAG-GNN(orange).
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/SHD_dependence_20_DAG_threshold0.3.pdf}
                \caption{SHD with dependent Gaussian}
                \label{fig:dep_gaussian_shd_20}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/FDR_dependence_20_DAG_threshold0.3.pdf}
                \caption{FDR with dependent Gaussian}
                \label{fig:dep_gaussian_fdr_20}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}{Dependent Noise(Cont'd)}
    Node size $m=30$ : our method(blue) performs better than DAG-GNN(orange).
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/SHD_dependence_30_DAG_threshold0.3.pdf}
                \caption{SHD with dependent Gaussian}
                \label{fig:dep_gaussian_shd_30}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/FDR_dependence_30_DAG_threshold0.3.pdf}
                \caption{FDR with dependent Gaussian}
                \label{fig:dep_gaussian_fdr_30}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}{Dependent Noise(Cont'd)}
    Node size $m=50$ : our method(blue) performs better than DAG-GNN(orange).
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/SHD_dependence_50_DAG_threshold0.3.pdf}
                \caption{SHD with dependent Gaussian}
                \label{fig:dep_gaussian_shd_50}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/FDR_dependence_50_DAG_threshold0.3.pdf}
                \caption{FDR with dependent Gaussian}
                \label{fig:dep_gaussian_fdr_50}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

>>>>

\documentclass{beamer}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{kotex}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage[backend=bibtex]{biblatex}
\addbibresource{references.bib}
\setbeamertemplate{bibliography item}{\insertbiblabel}
\setbeamertemplate{frametitle continuation}[from second][(cont'd)]

@@

\subsection{Graphical Model}

\begin{frame}[allowframebreaks]{DAG and SEM}
    \begin{itemize}
        \item Directed Acyclic Graph(DAG) $\mathcal G = (V, E)$ with $V = \{1, \dots, m\}$ is used to represent a causal structure \cite{pearl_2009}
        \item Linear Structural Equation Model(SEM) is defined as
        \begin{equation*}
            X_i = \sum_{j \in \text{Pa}(i)} \beta_{ij} X_j + \epsilon_i
        \end{equation*}
        where $\epsilon_i$ is a noise variable.
        \item Matrix form can be written as
        \begin{equation*}
            \mathbf X = \mathbf A^T \mathbf X + \mathbf \epsilon
        \end{equation*}
        where $\mathbf A$ is an adjacency matrix of $\mathcal G$ and $\mathbf \epsilon$ is a noise vector.
    \end{itemize}
    \framebreak
    \begin{itemize}
        \item If $(\mathbf I-\mathbf A^T)$ is nonsingular, SEM can be written as
        \begin{equation*}
            \mathbf X = (\mathbf I - \mathbf A^T)^{-1} \mathbf \epsilon
        \end{equation*}
        \item Usually we assume $\mathbf \epsilon \sim \mathcal N(\mathbf 0, \mathbf \Sigma)$ and
        \begin{equation*}
            \mathbf X \sim \mathcal N(\mathbf 0, (\mathbf I - \mathbf A^T)^{-1} \mathbf \Sigma (\mathbf I - \mathbf A)^{-1})
        \end{equation*}
        \item In most cases, we assume $\mathbf \Sigma = \mathbf I$ or $\mathbf \Sigma = \mathbf D$ where $\mathbf D$ is a diagonal matrix.
        \item DAG-GNN \cite{yu2019daggnn} generalizes linear SEM using a deep neural network.
        \begin{equation*}
            f_2^{-1}(X) = A^Tf_2^{-1}(X) + f_1(Z),
        \end{equation*}
    \end{itemize}
\end{frame}

\begin{frame}[allowframebreaks]{Semi-Markovian DAG}
    \begin{itemize}
        \item DAGs have many convenient properties (e.g. Markov property, d-separation, etc.)
        \item More general class of DAGs is acyclic directed mixed graphs (ADMGs) i.e. semi-Markovian DAGs
        \item Semi-Markovian DAG is a DAG where the noise variables have dependencies \cite{Shipster2006semimarkovian}
        \item The noise vector $\mathbf \epsilon$ is a semi-Markovian noise vector if
        \begin{equation*}
            \Sigma \in \mathcal{W}(G) := \{ \Sigma \in \mathbb R^{m\times m} : \Sigma_{ij} = \Sigma_{ji} = 0 \text{ if } i \notin \text{sib}(j) \}
        \end{equation*}
        where $\text{sib}(j) = \{i : i \sim j \text{ in } G\}$ \cite{wang2017empirical}
        \item Learn cyclic adjacency matrix \cite{wang2017empirical} or acyclic adjacency matrix
    \end{itemize}
\end{frame}

\subsection{DAG Learning}

\begin{frame}[allowframebreaks]{DAG Learning}
    \begin{itemize}
        \item DAG learning is a problem of estimating $\mathcal G$ from $\mathbf X$
        \item NP-hard problem \cite{ChickeringNPhard}
        \item Score-based methods \cite{Spirtes2000} and constraint-based methods \cite{ChickeringGES} are used
        \item NOTEARS\cite{zheng2018dags}: Combinatorial optimization to Continuous optimization problem
        \begin{equation*}
            \min_{\mathbf{A}} F(A) \text{ subject to } G(A) \in \mathbb{D} \Rightarrow 
            \min_{\mathbf{A}} F(A) \text{ subject to } h(A)=0
        \end{equation*}
        where $\mathbb D$ denotes the discrete space of DAGs on $m$ nodes and $h(A)=0$ iff $G(A)$ is acyclic.
    \end{itemize}

    \begin{block}{Theorem}
        A matrix $\mathbf A$ is the adjacency matrix of a DAG if and only if
        \begin{equation*}
            h(A) = \rm{tr}(\exp(\mathbf A \circ \mathbf A)) = m
        \end{equation*}
        where $\circ$ denotes the Hadamard product.
    \end{block}
    \begin{block}{Theorem}
        A matrix $\mathbf A$ is the adjacency matrix of a DAG if and only if
        \begin{equation*}
            \rm{tr}[(I + \alpha \mathbf A \circ \mathbf A)^m] = m
        \end{equation*}
        for any $\alpha>0$.
    \end{block}
\end{frame}

\subsection{VAE and Normalizing Flow}

\begin{frame}{VAE}
    \begin{itemize}
        \item Variational Autoencoder(VAE) is a generative model that learns a latent variable model
        \item It is trained by maximizing the evidence lower bound(ELBO)
        \begin{equation*}
            \mathcal L(\theta, \phi; \mathbf x) = \mathbb E_{q_\phi(\mathbf z | \mathbf x)}[\log p_\theta(\mathbf x | \mathbf z)] - \text{KL}(q_\phi(\mathbf z | \mathbf x) || p(\mathbf z))
        \end{equation*}
        \item $q_\phi(\mathbf z | \mathbf x)$ is an approximate posterior and $p_\theta(\mathbf x | \mathbf z)$ is a generative model
    \end{itemize}
\end{frame}

\begin{frame}{Normalizing Flow}
    \begin{itemize}
        \item Normalizing flow transforms a simple distribution to a complex distribution \cite{rezende15normalizingflows}
        \item It is a sequence of invertible transformations $f_t : \mathbb R^d \to \mathbb R^d$
        \begin{equation*}
            z_T = f_T \circ f_{T-1} \circ \cdots \circ f_1(z_0)
        \end{equation*}
        \item The probability density function of $z_T$ is
        \begin{equation*}
            p(z_T) = p(z_0) \left| \det \left( \frac{\partial f_T \circ f_{T-1} \circ \cdots \circ f_1(z_0)}{\partial z_0} \right) \right|
        \end{equation*}
    \end{itemize}
\end{frame}

\begin{frame}[allowframebreaks]{Linear IAF}
    \begin{itemize}
        \item Inverse Autoregressive Flow(IAF) is a normalizing flow that uses autoregressive transformation \cite{kingma2016iaf}
        \item Linear IAF is a special case of IAF where $f_t$ is a linear transformation
        \begin{equation*}
            \bf z_t = \bf \mu_t + \bf \sigma_t \odot \bf z_{t-1}.
        \end{equation*}
        \item Transformation of the linear IAF can be written as
        \begin{equation*}
            \mathbf{z}_T = \mathbf{L(x)} \cdot \mathbf{z}_0.
        \end{equation*}
        where $\mathbf{L(x)}$ is a lower triangular matrix from encoder network.
        \item We can use more flexible prior distribution by using normalizing flow.
        \item When $\mathbf z_0$ has diagonal covariance, $\mathbf z_T$ then has full covariance matrix.
        \item The KL divergence term of ELBO can be written as
        \begin{equation*}
            \begin{aligned}
                D_{KL}(q_\phi(\mathbf{z}_0 | \mathbf{x}) \Vert p(\mathbf{z}_T)) &=
                \log q_\phi(\mathbf{z}_0 | \mathbf{x}) - \log p(\mathbf{z}_T) \\
                &= -\frac{1}{2}(\mathbf{z}_0 -\mu_\phi)^T\Sigma_\phi^{-1}(\mathbf{z}_0 -\mu_\phi) + \frac{1}{2}\mathbf{z}_T^T\mathbf{z}_T \\
            \end{aligned}
        \end{equation*}
        where $\mu_\phi$ and $\Sigma_\phi$ are the mean and covariance matrix from the encoder network.
    \end{itemize}
\end{frame}

@@

\begin{frame}{Learning}
    \begin{itemize}
        \item The optimization procedure is minimizing the following loss function:
        \begin{equation*}
            \mathcal{L}(A, W, \lambda) =-\mathcal{L}_{\mathrm{ELBO}} + \tau \Vert A \Vert_1 + \lambda h(A) + \frac{c}{2} |h(A)|^2.
        \end{equation*}
        \item The second term is the L1 regularization term, which encourages sparsity of the adjacency matrix.
        \item $\tau=0.1 \times ({m\over 50})^2$ works well in our experiment setting.
        \item The third and fourth terms are the augmented Lagrangian terms.
        \item Gradually increasing the value of $c$ during the training process results the acyclicity constraint to zero.
    \end{itemize}
\end{frame}

@@

\begin{frame}[allowframebreaks]{Experiment}
    \begin{itemize}
        \item We compare our method with the DAG-GNN \cite{yu2019daggnn} with random graph datasets.
        \item thresholding value of extracting graph as 0.3
        \item Erdős–Rényi random graph with 5000 samples
        \item Node size : 10, 20, 30, and 50 nodes
        \item Noise type
        \begin{itemize}
            \item Independent : $\epsilon_i \sim \mathcal N(0, 1)$, $\epsilon_i \sim \text{Laplace}(0, 1)$, $\epsilon_i \sim \text{Exp}(1)$
            \item Dependent : $\epsilon_i \sim \mathcal N(0, \mathbf \Sigma)$ where $\mathbf \Sigma$ is a covariance matrix of semi-Markovian DAG
        \end{itemize}
    \end{itemize}
    \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{fig/mixed_graph_example.png}
        \caption{Example of mixed graph}
        \label{fig:mixed_graph_example}
    \end{figure}
    \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{fig/mixed_graph_adjacency.png}
        \caption{Adjacency matrix of previous example graph}
        \label{fig:mixed_graph_adjacency}
    \end{figure}
\end{frame}

@@

\begin{frame}[allowframebreaks]{Independent Noise}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/SHD_independence_normal.pdf}
                \caption{SHD with normal}
                \label{fig:ind_gaussian_shd}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/FDR_independence_normal.pdf}
                \caption{FDR with normal}
                \label{fig:ind_gaussian_fdr}
            \end{figure}
        \end{column}
    \end{columns}
    \begin{itemize}
        \item In the case of independent Gaussian noise, DAG-GNN(orange) performs better than our method(blue).
    \end{itemize}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/SHD_independence_laplace.pdf}
                \caption{SHD with Laplace}
                \label{fig:ind_laplace_shd}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/FDR_independence_laplace.pdf}
                \caption{FDR with Laplace}
                \label{fig:ind_laplace_fdr}
            \end{figure}
        \end{column}
    \end{columns}
    In the case of independent Laplace noise, our method(blue) performs better than DAG-GNN(orange).
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/SHD_independence_exponential.pdf}
                \caption{SHD with Exponential}
                \label{fig:ind_exponential_shd}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/FDR_independence_exponential.pdf}
                \caption{FDR with Exponential}
                \label{fig:ind_exponential_fdr}
            \end{figure}
        \end{column}
    \end{columns}
    In the case of independent Exponential noise, our method(blue) performs better than DAG-GNN(orange).

    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/Predicted Edges_independence_laplace.pdf}
                \caption{Number of predicted edges with Laplace}
                \label{fig:ind_laplace_edges}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/Predicted Edges_independence_exponential.pdf}
                \caption{Number of predicted edges with Exponential}
                \label{fig:ind_exponential_edges}
            \end{figure}
        \end{column}
    \end{columns}
    Considering the number of predicted edges with the SHD and FDR, our method(blue) performs better than DAG-GNN(orange).

    \begin{figure}
        \centering
        \includegraphics[height=3.5cm]{fig/comparison_indep_30_laplace_seed301.pdf}
        \label{fig:ind_exponential_comparison}
        \captionsetup{justification=centering}
        \caption{Comparison of estimated graphs with Laplace noise}
    \end{figure}
    \begin{figure}
        \centering
        \includegraphics[height=3.5cm]{fig/comparison_indep_30_exponential_seed301.pdf}
        \label{fig:ind_exponential_comparison}
        \captionsetup{justification=centering}
        \caption{Comparison of estimated graphs with Exponential noise}
    \end{figure}
\end{frame}

\begin{frame}[allowframebreaks]{Dependent Noise}
    Proportion of bidirectional edges fixed to 0.3.
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/SHD_dependence_30_DAG_threshold0.3.pdf}
                \caption{SHD with dependent Gaussian}
                \label{fig:dep_gaussian_shd}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/FDR_dependence_30_DAG_threshold0.3.pdf}
                \caption{FDR with dependent Gaussian}
                \label{fig:dep_gaussian_fdr}
            \end{figure}
        \end{column}
    \end{columns}

    Proportion of bidirectional edges fixed to 0.3.
    \begin{figure}
        \centering
        \includegraphics[height=4cm]{fig/Predicted Edges_dependence_30_DAG_threshold0.3.pdf}
        \caption{Number of predicted edges with dependent Gaussian}
        \label{fig:dep_gaussian_edges}
    \end{figure}
    \begin{itemize}
        \item In the case of dependent Gaussian noise, our method(blue) performs better than DAG-GNN(orange).
        \item Considering the stable number of predicted edges with the SHD and FDR, our method(blue) outperforms.
    \end{itemize}

    Node size $m=10$ : similar performance between DAG-GNN(orange) and our method(blue).
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/SHD_dependence_10_DAG_threshold0.3.pdf}
                \caption{SHD with dependent Gaussian}
                \label{fig:dep_gaussian_shd_10}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/FDR_dependence_10_DAG_threshold0.3.pdf}
                \caption{FDR with dependent Gaussian}
                \label{fig:dep_gaussian_fdr_10}
            \end{figure}
        \end{column}
    \end{columns}

    Node size $m=20$ : our method(blue) performs better than DAG-GNN(orange).
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/SHD_dependence_20_DAG_threshold0.3.pdf}
                \caption{SHD with dependent Gaussian}
                \label{fig:dep_gaussian_shd_20}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/FDR_dependence_20_DAG_threshold0.3.pdf}
                \caption{FDR with dependent Gaussian}
                \label{fig:dep_gaussian_fdr_20}
            \end{figure}
        \end{column}
    \end{columns}

    Node size $m=30$ : our method(blue) performs better than DAG-GNN(orange).
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/SHD_dependence_30_DAG_threshold0.3.pdf}
                \caption{SHD with dependent Gaussian}
                \label{fig:dep_gaussian_shd_30}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/FDR_dependence_30_DAG_threshold0.3.pdf}
                \caption{FDR with dependent Gaussian}
                \label{fig:dep_gaussian_fdr_30}
            \end{figure}
        \end{column}
    \end{columns}

    Node size $m=50$ : our method(blue) performs better than DAG-GNN(orange).
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/SHD_dependence_50_DAG_threshold0.3.pdf}
                \caption{SHD with dependent Gaussian}
                \label{fig:dep_gaussian_shd_50}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/FDR_dependence_50_DAG_threshold0.3.pdf}
                \caption{FDR with dependent Gaussian}
                \label{fig:dep_gaussian_fdr_50}
            \end{figure}
        \end{column}
    \end{columns}

    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[scale=0.3]{fig/Predicted Edges_dependence_10_DAG_threshold0.3.pdf}
                \caption{$m=10$}
                \label{fig:dep_gaussian_edges_10}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \centering
            \includegraphics[scale=0.3]{fig/Predicted Edges_dependence_20_DAG_threshold0.3.pdf}
            \caption{$m=20$}
            \label{fig:dep_gaussian_edges_20}
            \end{figure}
        \end{column}
    \end{columns}
    \begin{columns}
        \begin{column}{0.5\textwidth}
        \begin{figure}
            \centering
            \includegraphics[scale=0.3]{fig/Predicted Edges_dependence_30_DAG_threshold0.3.pdf}
            \caption{$m=30$}
            \label{fig:dep_gaussian_edges_30}
        \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
        \begin{figure}
            \centering
            \includegraphics[scale=0.3]{fig/Predicted Edges_dependence_50_DAG_threshold0.3.pdf}
            \caption{$m=50$}
            \label{fig:dep_gaussian_edges_50}
        \end{figure}
        \end{column}
    \end{columns}

    \begin{itemize}
        \item FPR : our method performs better than DAG-GNN.\\
        \item TPR : similar performance
    \end{itemize}
